### 2022-04-25 main.tex
Initial Overleaf Import

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Adversarial Penguin}
\author{Steve K Harris}
\date{April 2022}

\begin{document}

\maketitle

\section{Introduction}

\end{document}