### 2022-06-27 rebuttals/response-reviewer-1.md
formatting of response

# Independent Review Report, Reviewer 1

The article proposes a design pattern, Clinical Deployment Environments (CDE) with 05 principles (pillars). A design pattern is a general repeatable and implementable solution which is well demonstrated through the datasheet attached. My overall impression apart from a few specific comments is, that for each of the five principles (pillars), the problem description that they address may be clearly stated in the introduction section. The implementation example in the datasheet describes the first 02 pillars very clearly, however, an implementation example for the rest of the 03 pillars is missing. They should be included to prove the implementation. I would think that the Implementation science is a result of the first two pillars (i.e. RW-Dev and ML-Ops) and not essentially one of the pillars here. The explanation of 'Implementation Science' in the article also does not provide clear requirements for this pillar.

Other specific comments -
- The language style preferred for Frontiers in Digital Health Articles is US English while the article uses the GB English style. A revision is needed to match the editorial requirements
- A review of the state-of-the-art in the area may be included to highlight the gaps
- A thorough check for sentences, the flow of the information, and grammar will help increase the readability of the article. A few of the comments and errors are highlighted in the attached file.
- The diagram in Figure 1. does not match with the attached Image 1
- Suggest moving the figures as part of the datasheet where EMAP and FlowEHR are explained instead of keeping them in the main article
- An overall depiction of CDE and its pillars may be useful
- Datasheet, Page 2, Section 1.4.2, "... aggregation. (201, 2018) Flat (bulk) FHIR is not mature, nor designed for live interactions.(202, 2021)..." : Please update the references. They are not numbered in the bibliography and full information is missing.