### 2022-06-29 rebuttals/response-reviewer-2.md

# Independent Review Report, Reviewer 2

Thank you for the opportunity to review this manuscript from Harris, Bonnici, and colleagues regarding the concept of clinical deployment environments. The authors propose this concept as having potential to align development, deployment, and evaluation of electronic health record-based machine learning tools in the healthcare setting.

## Major Strengths
1. The pillars proposed by the authors have face validity and lend important terminology around important considerations within a relatively underdescribed area.

Thank you.

2. The figures are very nicely done and quite informative.

Thank you again.

## Major Limitations
1. The analogy to drug discovery is not a great fit, in my opinion, and its shortcomings reveal an opportunity to better-describe the relationships between ML4H tools and the problems they seek to solve.

- [ ] decide how to play this; see for other examples of using the same analogy "algorithmic stewardship"[@eaneff2020] esp b/c also talks about the team who should advice again making the argument for pillar 1; they also use the drug parallel (trials/hospital formulary/evaluations/stewardship)
> clinicians who are familiar with the language of data, patients, bioethicists, scientists, and safety and regulatory organizations.

2. Some areas of the prose could benefit from increased specificity and/or clarity.

- [ ] idea was to keep the main text narrative and then allow the interested reader to hit the ESM for details with the figures providing the bridge; possible compromise is to expand the captions for each so the flow of the text remains roughly the same??

3. Much of the supplemental content would be better served in the main text.

- [ ] ditto; decide how to play this

## Major Comments
1. The analogy to drug discovery is a poor fit for this essay. One key premise underlying drug discovery activities is that potentially-actionable targets have already been identified through prior research. For example, the decision to seek JAK2-interacting molecules emerged from the understanding that JAK2 regulates important cytokines in a number of disease states. The algorithmic equivalent of this kind of prior knowledge is not apparent to me. Without a clear understanding of (a) a well-articulated problem to be solved and (b) a reasonable mechanism/logic model in which an ML tool - provided it performs well - might affect stakeholder behavior, translational failure seems essentially guaranteed. Said differently, the bake-off between different ML approaches can really only be held once the model ("ligand")'s "receptor" has been mapped/described, which I would argue happens infrequently at present. As I read the authors' proposed pillars, I found myself reflecting on the need for this kind of formative work, informed by stakeholders, to ensure that the CDE can be used for properly-specified development, operations, and evaluation.

I might suggest omitting or significantly abbreviating this section, which could open space to insert valuable information from the supplement. I also think a paragraph commenting on the relationship of the CDE to the organization's "problem identification and specification" apparatus would be a very helpful "bookend" to the implementation and continuous evaluation portions of the essay.  
*We thank the reviewer for this comment but feel that changing the manuscript as they suggest would not be the right thing to do. The point of the drug development pipeline analogy is to articulate the complexity of the intellectual work required for innovation. No-one expects a drug to be developed from scratch in a 3-5 years. However, in many cases people are disappointed when healthcare AI algorithms do not yield practice-changing results in the timeframe of a single project. The benefit of the Drug Trial Phases is that people are easily able to understand the risk level, the size of trial and the appropriate endpoints. This same thinking needs to be related to algorithms. Similar points have been made by Pepe (2005) and Ferrante di Ruffano with respect to biomarker discovery.*
_To specifically address the JAK-2 example, this work would happen in the pre-clinical phase that we describe as algorithm discovery. The AI equivalent of JAK-2 (a targetable molecule) is a feature (a representation of data that has some utility in classification or prediction that also has a causal relationship with the phenomenon under study) _

2. Some areas of the prose could benefit from increased specificity and/or clarity. One of the challenges in the healthcare data science space is striking the right balance of technical language to meet a readership with different levels of familiarity to clinical medicine, coding/development, and operations. Some areas of the prose - e.g., lines 63-67 - skew a bit heavy into the data science jargon and might benefit from additional explanations or a bit less jargon. It may be worth the authors having a colleague with less technical familiarity provide some informal copy-editing to help with readability for an audience of varying levels of expertise.

Thank you. This was helpful. I have asked two colleagues to separately proof the manuscript. They have made a series of small stylistic changes that I have incorporated into the manuscript.

e.g. lines 63-67

Before:

> Given the sensitive nature of health data, these initiatives depend on expanding investment into Trusted Research Environments (TRE). [@2021c] Data flows from source (primary, secondary, social care and elsewhere) to a single secure landing zone where research teams write the code to link, clean and analyse the data. The insights return to the bedside through clinical guidelines and policy. This offline '_data-to-code_' approach is also the dominant design pattern in ML4H projects but is fundamentally flawed.

After

> Given the sensitive nature of health data, these programmes depend on Data Safe Havens (also known as Trusted Research Environments [TRE]). [@2021c] TREs are an example of 'data-to-modeler' (DTM) designs where source data (from primary, secondary, and social care) **moves to** a separate, secure landing zone. Here research teams link, clean and analyse the data. Their insights eventually return to the bedside through clinical guidelines and policy. In ML4H, DTM is also the dominant design pattern but is fundamentally flawed.

A detailed diff of these changes is available [here](https://github.com/docsteveharris/2022-adversarial-penguin/pull/7)

3. The supplement - especially items 1 and 2 - are very nice and highly practical. Some elements of these items could serve to address issue 2 above. Further, the ML-Ops design principles are particularly helpful explicating how code-to-data can address the limitations of data-to-code. I think it would be great if these could be a bit more prominent, ideally in the main manuscript.

Minor Comments
1. The manuscript has a number of instances with unspecified pronouns, e.g., sentences starting with "This" or "That," which do not specify what "This" or "That" actually are.

Thank you. We hope we have corrected this with the proof reading and recommendations (esp. reviewer 1).

2. Supplement page 2, typo in the ML-0PS header (zero instead of O)

Thanks for this careful proofread. I double checked. I think this is a quirk of the font that Overleaf is using for the online submission and I can confirm the source document uses `O` not `0`.

i.e.

```markdown
# ML-Ops (for health)
```

not

```markdown
# ML-0ps (for health)
```